---
title: "County usa-facts confirmed case corrections with backfilling"
author: "Delphi Group - Last run:"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
params:
  window_size: 14
  start_date: NULL
  sig_cut: 3
  size_cut: 20
  sig_consec: 2.25
  outlier_start_date: "2020-03-01"
  time_value_flag_date: !r Sys.Date()
  cache_data: FALSE
  backfill_lag: 30
  excess_cut: 0
  write_RDS: TRUE
  covidcast_data_source: "usa-facts"
  covidcast_signal: "confirmed_incidence_num"
  corrections_db_path: "./data_corrections-mp.sqlite"
  editor_options: 
  chunk_output_type: inline
---

```{r, message=FALSE, warning=FALSE}
library(covidcast)
library(dplyr)
library(tidyr)
library(lubridate)
library(RcppRoll)
library(cowplot)
library(ggplot2)
library(DT)
library(ggforce)
knitr::opts_chunk$set(warning = FALSE, message=FALSE, dev="CairoSVG")
source("process-funs.R")
source("corrections.R")
attach(params)
```

Data retrieval
```{r, cache=params$cache_data}
counties <-  suppressMessages(
  covidcast_signal(
    covidcast_data_source, covidcast_signal,
    geo_type = "county", 
    start_day = outlier_start_date)
)
```


Data Wrangling notes:

* Only counties with > 30 days' available data will be selected
* Counties whose maximum daily_confirmed_case <10 are filtered out
* Top 300 of counties sorted by total cases will be corrected

```{r}
todump <- counties %>% group_by(geo_value) %>% 
  summarise(
    ava_value_count = sum(!is.na(value)),
    case_sum = sum(value,na.rm = T),
    max_case = max(value)
    ) %>% 
  filter(max_case>=10, 
         ava_value_count>=30, 
         as.numeric(geo_value) %% 1000 > 0) %>% 
  arrange(desc(case_sum)) %>%
  top_n(300, wt=case_sum) %>%
  select(geo_value) %>%
  pull()
county_filtered <- filter(counties, geo_value %in% !!todump)
```





```{r calculate-roll-stats}
county_filtered <- county_filtered %>% group_by(geo_value)  %>% mutate(
  fmean = roll_meanr(value, window_size),
  # smean = roll_mean(value, window_size, fill = NA),
  fmedian = roll_medianr(value, window_size),
  smedian = roll_median(value, window_size, fill = NA),
  fsd = roll_sdr(value, window_size),
  ssd = roll_sd(value, window_size,fill = NA),
  fmad = roll_medianr(abs(value-fmedian), window_size,na.rm=TRUE),
  smad = roll_median(abs(value-smedian), na.rm=TRUE),
  ftstat = abs(value-fmedian)/fsd, # mad in denominator is wrong scale, 
  ststat = abs(value-smedian)/ssd, # basically results in all the data flagged
  flag = 
    (abs(value) > size_cut & !is.na(ststat) & ststat > sig_cut ) | # best case
    (is.na(ststat) & abs(value) > size_cut & !is.na(ftstat) & ftstat > sig_cut) | 
      # use filter if smoother is missing
    (value < -size_cut & !is.na(ststat) & !is.na(ftstat)), # big negative
    #(fmean > 10 & fmean< 20 & value > 2*sig_cut*fmean)
  flag = flag | # these allow smaller values to also be outliers if they are consecutive
    (dplyr::lead(flag) & !is.na(ststat) & ststat > sig_consec) | 
    (dplyr::lag(flag) & !is.na(ststat) & ststat > sig_consec) |
    (dplyr::lead(flag) & is.na(ststat) & ftstat > sig_consec) |
    (dplyr::lag(flag) & is.na(ststat) & ftstat > sig_consec),
  # RI_daily_flag = (geo_value=="44007" & value!=0 & 
  #                    lead(value,n=1L)!=0 & lead(value,n=2L)!=0 
  #                  & lead(value,n=3L)!=0)
  ##flag = flag & (time_value < ymd("2020-11-01") | value < -size_cut)
  flag = flag & 
    (time_value < ymd(time_value_flag_date) | value < -size_cut),
  flag = flag | 
    (time_value == "2020-11-20" & as.numeric(geo_value) %/% 1000 == 22)
  #Louisiana backlog drop https://ldh.la.gov/index.cfm/newsroom/detail/5891
  )

county_filtered <-  county_filtered %>% 
  mutate(FIP = substr(geo_value,1,2)) %>% 
  mutate(state = names(STATE_TO_FIPS)[match(FIP,STATE_TO_FIPS)]) %>% 
  select(-FIP) %>% relocate(state,.after=geo_value)
```




## Make corrections

Now we use the "multinomial" smoother to backfill the excess of any flagged outliers. Some notes:

* We use a new function `corrections_multinom_roll()` to do the backfill.
* It backfills deterministicly rather than randomly.
* It rounds alternate days up or down to try to avoid too much integers such that the sum is the excess.
* Optionally allows for filling non-uniformly.


```{r, eval=FALSE}
# All the counties treated the same
corrected_counties_old <- county_filtered %>% mutate(
   # FIPS = as.numeric(geo_value),
   # excess = value,
   excess = value - na_replace(smedian, fmedian),
   excess = floor(excess - excess_cut*sign(excess)*na_replace(smad,fmad)),
   corrected = corrections_multinom_roll(
     value, excess, flag, time_value, backfill_lag,
     reweight=function(x) exp_w(x, backfill_lag)))
```




```{r}
# RI reports only weekly
corrected_counties <- county_filtered %>% 
  mutate(
    # FIPS = as.numeric(geo_value),
    excess = value - na_replace(smedian, fmedian),
    excess = floor(excess - excess_cut*sign(excess)*na_replace(smad,fmad)),
    flag_bad_RI = (state == "RI"  & value > 10 & lag(value) == 0),
    corrected = corrections_multinom_roll(
      value, value, flag_bad_RI, time_value, 7),
    # flag_bad_VA =(geo_value == "51059"  && lag(value) > 0),
    # corrected = corrections_multinom_roll(
    #   value, value, flag_bad_VA, time_value, FIPS, 7),
    corrected = corrections_multinom_roll(
      corrected, excess, (flag & !flag_bad_RI), time_value, 
      backfill_lag, 
      reweight=function(x) exp_w(x, backfill_lag)),
    corrected = evalforecast:::multinomial_roll_sum(corrected),
    corrected = corrected + # imputes forward due to weekly releases
      missing_future(TRUE, time_value, value, fmean)
    )
```



## Visualize corrected counties

```{r show-corrections, fig.height = 80, fig.width = 20, dev="CairoSVG"}
simple_labs = covidcast::fips_to_name(unique(corrected_counties$geo_value))
simple_labs = gsub(" County| city| Parish", "", simple_labs)
sta = names(STATE_TO_FIPS)[match(substr(names(simple_labs), 1, 2), STATE_TO_FIPS)]
nn = names(simple_labs)
simple_labs = paste(simple_labs, sta)
names(simple_labs) = nn

mtv <- max(corrected_counties$time_value)
corrected_counties %>% group_by(geo_value) %>% 
  filter(any(flag == 'TRUE')) %>% 
  dplyr::select(geo_value, time_value, value, corrected, flag) %>%
  pivot_longer(value:corrected) %>%
  ggplot(aes(time_value))+geom_line(aes(y=value, color=name))+
  geom_point(
    data = filter(corrected_counties, flag), 
    aes(y=value), color="red")+ 
  facet_wrap(~geo_value, scales = "free_y", ncol = 5,
             labeller = labeller(geo_value = simple_labs))+
  theme_cowplot() + xlab("date") +
  coord_cartesian(xlim = mtv + c(-90,0)) +
  scale_color_viridis_d(end=.5) + 
  scale_x_date(date_breaks = "months" , date_labels = "%b")
```


## Show all corrected time points

```{r}
sum_check = corrected_counties %>% 
  summarise(original = sum(value, na.rm=TRUE),
            corrected = sum(corrected, na.rm = TRUE)) %>% ungroup() %>%
  mutate(diffs = abs(original-corrected)) %>%
  filter(diffs > 1e-8) %>%
  select(geo_value, diffs)

sum_check

tosave <- corrected_counties %>% 
  mutate(output = abs(corrected - value) > 0) %>%
  filter(output) %>% ungroup() %>%
  dplyr::select(geo_value, time_value, value, corrected) %>%
  transmute(
    location=as.character(geo_value),
    location_name=as.character(NA),
    reference_date=as.Date(time_value), 
    issue_date=as.Date(NA),
    variable_name = paste(covidcast_data_source, covidcast_signal, sep = "_"),
    value = as.double(value),
    new_value = as.double(corrected),
    correction_date = Sys.Date(),
    description = as.character("")
    ) %>%
  filter(reference_date >= params$outlier_start_date) %>%
  arrange(reference_date)


corrected_counties %>% 
  mutate(output = abs(corrected - value) > 0) %>%
  filter(output) %>%
  dplyr::select(geo_value, time_value, value, corrected) %>%
  transmute(time_value=time_value, 
            geo_value=geo_value,
            orig_value = value,
            replacement = corrected
            ) %>%
  datatable(
    options = list(scrollX = TRUE, scrollY = "300px",paging = FALSE),
    rownames = NULL) 
```




```{r eval=params$write_RDS}
update_corrections(corrections_db_path, "county", tosave)
```
